---
title: "PML Assign1"
output: html_document
---



```{r warning=FALSE}
library(caret)
trdata <- read.csv("pml-training.csv",header = TRUE,sep = ",", na.strings=c("#DIV/0!") )
for(i in c(8:ncol(trdata)-1)) {trdata[,i] = as.numeric(as.character(trdata[,i]))}
trdata <- trdata[,colSums(is.na(trdata)) == 0]
trdata$cvtd_timestamp = as.POSIXct(strptime(trdata$cvtd_timestamp, "%d/%m/%Y %H:%M"))
trdata <- subset(trdata, select = -c(user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2))
```

## Data Process Steps:
na.strings used to set missings values, blank fields, div by zero in numeric columns as NA.
Convert all non numeric columns to numeric except Classe.
Further remove columns with NA as they inhibit creating model

Feature extracted every .5 to 2.5 sec: Euler angles(roll, pitch, yaw), raw accelerometer, gyroscope and magnetometer readings.
For Euler angles eight features were calculated: mean, variance, sd, max, min, amp, kurtosis, skewness.
Total features: 96.
Time could be a useful predictor of classe. So is index in case a variable is not factored.
Username, window, timestamp are not interpretable predictor.

## Repeat data processing steps for test data:

```{r warning=FALSE}
tsdata <- read.csv("pml-testing.csv",header = TRUE,sep = ",", na.strings=c("#DIV/0!") )
for(i in c(8:ncol(tsdata)-1)) {tsdata[,i] = as.numeric(as.character(tsdata[,i]))}
tsdata <- tsdata[,colSums(is.na(tsdata)) == 0]
tsdata$cvtd_timestamp = as.POSIXct(strptime(tsdata$cvtd_timestamp, "%d/%m/%Y %H:%M"))
tsdata <- subset(tsdata, select = -c(user_name, new_window, num_window, raw_timestamp_part_1, raw_timestamp_part_2))
```
## How you used cross validation

```{r}
inTrain <- createDataPartition(trdata$classe, p = .7, list = F)
training <- trdata[inTrain,]
testing <- trdata[-inTrain,]
```

## The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set.

To minimise error and exploit the varacity of trees we choose boosting with trees(gbm) as the method for prediction.
Control the computational nuances of the train function using resampling method cv.
```{r}
ctrl <- trainControl(allowParallel=T, method="cv", number=4)
modFit <- train(classe ~ ., method="gbm",data=training,verbose=FALSE, trControl=ctrl)
print(modFit)
predcl <- predict(modFit, newdata = testing)

sum(predcl == testing$classe) / length(predcl)
```


As evident from the the summary of fit, this is a good model.



Citation:
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013. Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3BKAD0M8G
